---
title: "week 4 hw"
author: "Madison Dunn"
date: "9/30/2021"
output: github_document
---

Github repo: https://github.com/madisondunn8/433-week4hw

```{r}
library(nycflights13)
library(ggplot2)
library(dplyr)
```

What time of day should you fly if you want to avoid delays as much as possible? Does this choice depend on anything? Season? Weather? Airport? Airline? Find three patterns (“null results” are ok!). Write your results into Rmarkdown. Include a short introduction that summarizes the three results. Then, have a section for each finding. Support each finding with data summaries and visualizations. Include your code when necessary. This shouldn’t be long, but it might take some time to find the things you want to talk about and lay them out in an orderly way.

# Introduction

In general, it appears that you should fly earlier in the day if you want to avoid delays as much as possible. This makes sense intuively, as once flights begin getting delayed, flights that occur after those delayed flights will also likely get delayed - somewhat of a domino effect. After investigating the flights data more, it appears that other variables have a relationship with when to fly to avoid delays, including carrier, month of the year, and weather patterns, specifically average visibility. 

```{r}
flights %>%
  group_by(hour) %>%
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(avg_arr_delay) %>% 
  ggplot(aes(hour, avg_arr_delay)) +
  geom_jitter()
```
  
# First Finding

The first pattern I noticed was between the time of average arrival delay and flight carrier. In general, you should still fly earlier in the day because as the day goes on, average arrival delays increase. However, some carriers tend to have a different effect on when to fly. For example, avoid carriers FL, B6, and EV, especially in the later half of the day because they have higher average arrival delays - almost all the data points are positive, which means on average they do have arrival delays.. Carriers VX, US, and AA in the first half of the day tend to arrive earlier than expected, since they have negative average arrival delays. These carriers are better to fly through to avoid arrival delays. 

```{r}
flights %>%
  group_by(hour, carrier) %>%
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(avg_arr_delay) %>% 
  ggplot(aes(x=hour, y = avg_arr_delay)) +
  geom_point()+
    facet_wrap(~carrier)+
  ylim(NA,40) +
  geom_hline(yintercept=0)
```

# Second finding 

Another pattern I noticed was the relationship between time of day for average arrival delays and month of year. Flights around the holidays (December = 12, January = 1) seem to have longer average arrival delays than flights during the rest of the year. Therefore month of the year could affect delays. You should fly in the mornings not during the holiday months if possible to avoid delays. Flying early in the day in the fall, such as September (9) and October (10), flights tend to have negative average arrival delays, meaning they tend to arrive early rather than be delayed. 


```{r}
flights %>%
  group_by(hour, month) %>%
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(avg_arr_delay) %>% 
  ggplot(aes(x=hour, y = avg_arr_delay)) +
  geom_jitter()+
    facet_wrap(~month) +
 scale_y_continuous() +
    ylim(NA,40) +
    geom_hline(yintercept=0)
    
```

# Third finding

The third finding I noticed was the relationship between weather, specifically visibility, and time of day for average arrival delay of flights. In general, the pattern appears to persist between time of day and average arrival day. As the day goes on, average arrival delay increases as well. However, through looking at weather patterns, specifically average visability, flights with higher average visability tend to have a lower average arrival time. This can be seen by the more light blue colored dots, which correlate to high average visibility. Darker colored dots, such as those above 75 minute average arrival delates, tend to have much lower average visability. As though day goes on, there is still a slight upward pattern of average arrival delay. Therefore to avoid delays, you would want to fly on days with higher visibility, in the morning or earlier part of the day. 


```{r}
left_join(flights, weather) %>%
  group_by(visib, hour) %>%
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE),avg_visib=mean(visib)) %>%
  arrange(avg_arr_delay) %>% 
  ggplot(aes(hour, avg_arr_delay)) +
    geom_jitter(aes(color = avg_visib), size=2)
```





